---
title: "group_midus"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Including Code

You can include R code in the document as follows:

```{r, include=FALSE}
source("R/my_read_rda.R", local = knitr::knit_global())
source("R/filter_lipid.R")
source("R/filter_confounders.R")
source("R/filter_pathway.R")
source("R/filter_optimism.R")
source("R/fig1plot.R")
source("R/gen_table1_lipid.R")
```

## Introduction
In this Rmarkdown file we will attempt to reproduce the figures, tables and analyses presented in the paper [_Relation between Optimism and Lipids in Midlife_.](https://www.ajconline.org/article/S0002-9149(13)00388-3/pdf)

1. Boehm, J. K., Williams, D. R., Rimm, E. B., Ryff, C., & Kubzansky, L. D. (2013). Relation between Optimism and Lipids in Midlife. The American Journal of Cardiology, 111(10), 1425-1431.
http://doi.org/10.1016/j.amjcard.2013.01.292

In 1995, MIDUS survey data were collected from a total of 7,108 participants. The baseline sample was comprised of individuals from four subsamples: (1) a national RDD (random digit dialing) sample ($n = 3,487$); (2) oversamples from five metropolitan areas in the U.S. ($n = 757$); (3) siblings of individuals from the RDD sample ($n = 950$); and (4) a national RDD sample of twin pairs ($n = 1,914$). All eligible participants were non-institutionalized, English-speaking adults in the contiguous United States, aged 25 to 74. All respondents were invited to participate in a phone interview of approximately 30 minutes in length and complete 2 self-administered questionnaires (SAQs), each of approximately 45 pages in length. In addition, the twin subsample was administered a short screener to assess zygosity and other twin-specific information. With funding provided by the National Institute on Aging, a longitudinal follow-up of MIDUS I began in 2004. Every attempt was made to contact all original respondents and invite them to participate in a second wave of data collection. Of the 7,108 participants in MIDUS I, 4,963 were successfully contacted to participate in another phone interview of about 30 minutes in length. MIDUS II also included two self-administered questionnaires (SAQs), each of about 55 pages in length, which were mailed to participants. The overall response rate for the SAQs was 81\%. Over 1,000 journal articles have been written using MIDUS I and II data since 1995.

Here we attempt to reproduce the findings of [1] and critique the reproducibility of the article. This particular article focuses only on MIDUS II data, including biomarker data, and investigates the relationship between optimism and lipids. The MIDUS II data and supporting codebook and other documents can be downloaded \href{https://www.icpsr.umich.edu/icpsrweb/ICPSR/studies/4652?archive=ICPSR&q=MIDUS+}{here}. The data can be downloaded in multiple formats. Here we use the R data files and perform all data cleaning and analyses in R version 3.5.1. The biomarker data can be downloaded \href{https://www.icpsr.umich.edu/icpsrweb/NACDA/studies/29282}{here}.

## Data Dictionary
This manuscript uses several variables from multiple data files. Some of these variables don't have intuitive names and need to be manually looked up either online or in the codebooks provided in the data downloads. We generated a data dictionary to our understanding of the naming conventions.

Load packages
```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyverse)
library(ggplot2)
```

We are trying to keep all functions well documented. This command allows us to have
package-like documentation for all of the functions.

```{r}
if (!require('devtools')) install.packages('devtools')
devtools::document()
```

## Read data

First we load the data. 29282-0001-Data contains the analysis-associated data, 
while 04652-0001-Data contains the midus clinical data. 
```{r}
getwd()
load('data/29282-0001-Data.rda')
load('data/04652-0001-Data.rda')
```

We have to merge the two tables based on the MIDUS II ID number

```{r}
data = inner_join(da04652.0001, da29282.0001, by = c("M2ID", "M2FAMNUM"),suffix = c('','.2'))
print(dim(data))
```

Data has 1054 rows at the beginning after merging the two tables. 
Now we are going to try and reproduce the preprocessing steps such that we can 
obtain the 990 individuals they used for the paper analysis.

## Wrangle data

### Step 0. Filter optimism variables

Optimism is assessed using the 6-item Life-Orientation test. In the codebook we have
found that the B1SORIEN column contains the test value, already processed as 
explained in the paper while columns ('B1SORIEN','B1SE10A', 'B1SE10B', 'B1SE10C', 'B1SE10D', 'B1SE10E','B1SE10F')
contain the single item scores.

Here we are filtering the rows to remove the individuals
who do not have an optimism score.

```{r}
data_after_fo = data %>% filter_optimism()
nrow(data_after_fo)
```

We are left with 1050 samples. 
However, by visually inspecting the table we can see that for some patients, 
although we have a B1SORIEN value, some of the items are reported as NAs. 
This seems weird, we do not understand how they computed the final score, 
hence we proceed to remove also these patients.

```{r}

optimism_columns <- c('B1SORIEN','B1SE10A', 'B1SE10B', 'B1SE10C', 'B1SE10D', 'B1SE10E','B1SE10F')
data_after_fo2 <- data_after_fo %>%
    drop_na(any_of(optimism_columns))
nrow(data_after_fo2)
```

We are left with 1041 samples.

Secondly, we clean the columns relative to lipids (Total cholesterol, HDL, LDL, triglicerydes).

### Step 1 Filter lipid measurements
We will focus on 24 columns of our interests and drop rows with missing values in "B4BCHOL", "B4BTRIGL", "B4BHDL" and "B4BLDL" columns.
```{r}
lipid_columns = c("B4BCHOL", "B4BTRIGL", "B4BHDL", "B4BLDL")
data_after_fl = data_after_fo2 %>% filter_lipid()
nrow(data_after_fl)
```

After filtering lipid measurements, we have 1030 rows left

We also have to clean and filter the pathway variables which are those relative to lifestyle
and concern drinking, smoking and diet.

### Step 2 Filter pathway variables
```{r}

data_after_fp = filter_pathway(data_after_fl)
nrow(data_after_fp)
pathway_columns = c("B4PBMI", "alcohol_consumption", "smoking_status", "reg_exercise","score_sum")
```

We are left with 1020 samples. This filtering was tricky because we found that
many choices are not adequately explained in the paper: what columns are used for 
the drinking habits? 

### Step 3 Filter potential confounders
Finally we filter the potential confounders, such as age, sex, income..

```{r}
confounder_columns = c('B1PB1','B1STINC1','B4ZB1SLG','B1SCHROX','B4H26','B4H33','B4H25','B4PBMI','B1SNEGAF')

data_after_fc = filter_confounders(data_after_fp, confounder_columns)
nrow(data_after_fc)
```

Here we are left with 994 individuals, 4 more than those reported in the paper.

Although we are not able to completely reproduce the data wrangling steps, 
we are now going to reproduce some of the figures and tables and assess whether the 
general trends reported in the paper are robust to this slight change in the population 
make-up.

Let's keep only the columns of interest
```{r}
all_columns = c(optimism_columns, lipid_columns, pathway_columns, confounder_columns)
View(data_after_fc[,all_columns])
```



## Figure 1

First, we attempt to reproduce Figure 1. 
Figure 1 shows the frequency distribution of 990 optimism scores 
(mean +- SD: 23.95 +- 4.69), 
with black representing the lowest tertile of optimism (6 to 22), 
gray, middle tertile of optimism (23 to 26), 
and white, highest tertile of optimism (27 to 30)

In our data the average and standard deviation might be slightly different
```{r}
mean(data_after_fc$B1SORIEN)
sd(data_after_fc$B1SORIEN)
```

Indeed we have a slight deviation in mean, that is now 23.91 vs the 23.05 reported in the table. 
Standard deviation is instead consistent with what was reported.


```{r}
gen_fig1(data_after_fc)
```
The histogram of the optimism scores seems to be slightly different, for instance 
there seem to be more samples with score 22.

## Table 1

We then proceed to reproduce table 1. We are gonna split it in different chunks, 
based on the lipid/confounder/pathway groups.

```{r}
gen_table1_lipid(data_after_fc)
```

